One-arm spirals in locally isothermal, radially structured self-gravitating discs

### 2014-11-05 paper/discussion.tex
minor updates

\subsection{Role of self-gravity and disc structure} 
We note that the background torque density $T_\mathrm{BG}$ does not depend on  
self-gravity, and we only considered a self-gravitating perturbation to
obtain an appropriate wavenumber $k$ needed to evaluate
Eq. \ref{theoretical_rate}. Thus, self-gravity is not directly
responsible for instability. However, 
our numerical simulations show that the $m=1$ spiral is
confined between two $Q$-barriers, where real solutions to the local
dispersion relation is possible (Eq. \ref{wavenumber}), as a result of the adopted disc
structure. Thus, the role of disc structure and self-gravity is to
allow a neutral mode to be set up and confined, which is then
destabilised through the background torque density.

@@

>>>>

\subsection{Role of self-gravity and disc structure} 
We note that the background torque density $T_\mathrm{BG}$ does not depend on  
self-gravity, and we only considered a self-gravitating perturbation to
obtain an appropriate wavenumber $k$ needed to evaluate
Eq. \ref{theoretical_rate}. Thus, self-gravity is not directly
responsible for instability. However, 
our numerical simulations show that the $m=1$ spiral is
confined between two $Q$-barriers, where real solutions to the local
dispersion relation is possible (Eq. \ref{wavenumber}), as a result of the adopted disc
structure. Thus, the main role of disc structure and self-gravity is to
allow a neutral mode to be set up and confined, which is then
destabilised through the background torque density.

@@

\subsubsection{The confined spiral as an external potential}
One possible effect of self-gravity is to allow the one-arm spiral 
to act as an external potential for the wider disc. This effect is
analogous to disc-satellite interaction
\citep{goldreich79}, and may lead to instability 
if the angular momentum associated with the spiral has the opposite 
sign to that of the density waves it induces in the exterior disc
\citep{lin11a}.

Considering the spiral as a external potential (c.f. an embedded
satellite), we expect it to  exert a torque on the disc at Lindblad
and co-rotation resonances.

%speculations - long term balance. forced temp gradient drives
%instability (disc gains ang mom, spiral loses ang mom). but if spiral
%shocks and dissiple, it dumps negative angular momentum (disc gains
%negative ang mom). balance between gain from forced temp gradient and
%loss due to dissiplatoin

### 2014-11-05 paper/results.tex
minor updates

\subsection{Dependence on the imposed temperature gradient} 
At this point we hypothesize the growth of the $m=1$ spiral mode is
associated with the imposed temperature gradient. The torque
responsible for its growth, Eq. \ref{baroclinic_torque}, is
proportional to $q$, the power-law index used to set the sound-speed
profile (Eq. \ref{sound-speed}). To test this, we carried out
simulations with $q\in[0,1]$ using only an $m=1$ initial 
perturbation. However, to keep the Toomre profiles similar, we adjust
the surface density power-law index such that $s = (3+q)/2$ for each
case presented below. %should not be important, local effects

@@

>>>>

\subsection{Dependence on the imposed temperature profile} 
At this point we hypothesize the growth of the $m=1$ spiral mode is
associated with the imposed temperature gradient. The torque
responsible for its growth, Eq. \ref{baroclinic_torque}, is
proportional to $q$, the power-law index used to set the sound-speed
profile (Eq. \ref{sound-speed}). To test this, we carried out
simulations with $q\in[0,1]$ using only an $m=1$ initial 
perturbation. However, to keep the Toomre profiles similar, we adjust
the surface density power-law index such that $s = (3+q)/2$ for each
case presented below. %should not be important, local effects

@@

We also performed a series of simulations with aspect-ratio
$h\in[0.03,0.07]$ with fixed $q=1$. This also affects the temperature
gradient since $c_s \propto h$. However, with other parameters equal
to that in the fiducial simulation,  varying $h$ also changes the disc
mass. For $h\in[0.03,0.07]$ the total disc mass ranges from
$M_d=0.052M_*$ to $M_d=0.12M_*$ and the 
mass within $R\in[R_1,R_2]$  ranges from $0.033M_*$ to $0.062M_*$.

Fig. \ref{fargo_varh_growth} shows the growth rates of the $m=1$
spiral in $R\in[R_1,R_2]$ as a function of $h$. Growth rates increases
with $h$, roughly as 
\begin{align}
  \gamma \simeq 0.10h\Omega_k(R_0).  
\end{align}
However, the linear fit is less good than for variable $q$ above. This
may be due to the change in the total disc mass when $h$ changes. We
find no qualitative difference between the spiral pattern that
emerges.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_ratemax_fargo_varh.ps}    
  \caption{Growth rates of the $m=1$ spiral mode as a function of the
    disc aspect-ratio $h$ (asterisks). A linear fit is also
    plotted (dotted line). 
    \label{fargo_varh_growth}} 
\end{figure}

### 2014-11-05 paper/results_3d.tex
minor updates

\section{Three-dimensional simulations}
In this section we briefly review results from 3D simulations carried 
out using the ZEUS-MP and PLUTO codes. The main purpose is to check
verify the above results with different numerical codes, and validate 
the 2D approximation.

>>>>

\section{Three-dimensional simulations}
In this section we briefly review results from 3D simulations carried 
out using the ZEUS-MP and PLUTO. The main purpose is to check 
the above results with different numerical codes, and validate  
the 2D approximation.
